---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# JMH

<!-- badges: start -->
[![R-CMD-check](https://github.com/shanpengli/JMH/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shanpengli/JMH/actions/workflows/R-CMD-check.yaml)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/JMH)](https://cran.r-project.org/package=JMH)
[![](https://cranlogs.r-pkg.org/badges/grand-total/JMH)](https://cran.r-project.org/package=JMH)
[![CRAN_time_from_release](https://www.r-pkg.org/badges/ago/JMH)](https://cran.r-project.org/package=JMH)
[![CRAN_Status_Badge_version_last_release](https://www.r-pkg.org/badges/version-last-release/JMH)](https://cran.r-project.org/package=JMH)
<!-- badges: end -->

The `JMH` package jointly models both mean trajectory and within-subject variability of the longitudinal biomarker together with the (competing risks) survival outcome.

## Installation

You can install the development version of JMH from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("shanpengli/JMH")
```

# Example

The `JMH` package comes with several simulated datasets. To fit a joint model, we use `JMMLSM` function. 

```{r example}
library(JMH)
data(ydata)
data(cdata)
## fit a joint model
fit <- JMMLSM(cdata = cdata, ydata = ydata, 
              long.formula = Y ~ Z1 + Z2 + Z3 + time,
              surv.formula = Surv(survtime, cmprsk) ~ var1 + var2 + var3,
              variance.formula = ~ Z1 + Z2 + Z3 + time, 
              quadpoint = 10, random = ~ 1|ID, print.para = FALSE)
fit
```


The `JMH` package can make dynamic prediction given the longitudinal history information. Below is a toy example for competing risks data. Conditional cumulative incidence probabilities for each failure will be presented.

```{r, eval=TRUE}
cnewdata <- cdata[cdata$ID %in% c(122, 152), ]
ynewdata <- ydata[ydata$ID %in% c(122, 152), ]
survfit <- survfitJMMLSM(fit, seed = 100, ynewdata = ynewdata, cnewdata = cnewdata, 
                         u = seq(5.2, 7.2, by = 0.5), Last.time = "survtime",
                         obs.time = "time", method = "GH")
survfit
oldpar <- par(mfrow = c(2, 2), mar = c(5, 4, 4, 4))
plot(survfit, include.y = TRUE)
par(oldpar)
```

To assess the prediction accuracy of the fitted joint model, we may run `PEJMMLSM` to calculate the Brier score. 

```{r, eval=TRUE}
PE <- PEJMMLSM(fit, seed = 100, landmark.time = 3, horizon.time = c(4:6),
               obs.time = "time", method = "GH", 
               n.cv = 3)
summary(PE, error = "Brier")
```

An alternative to assess the prediction accuracy is to run `MAEQJMMLSM` to calculate the prediction error by comparing the predicted and empirical risks stratified on different risk groups based on quantile of the predicted risks.

```{r, eval=TRUE}
## evaluate prediction accuracy of fitted joint model using cross-validated mean absolute prediction error
MAEQ <- MAEQJMMLSM(fit, seed = 100, landmark.time = 3, 
                   horizon.time = c(4:6), 
                   obs.time = "time", method = "GH", n.cv = 3)
summary(MAEQ)
```
